Contextual recruitment of selective attention can be updated via changes in task-relevance

### 2018-12-09 redoc_manuscript.Rmd
first draft

output:
 # papaja::apa6_word: default
  papaja::apa6_pdf: default
  #redoc::rdocx_reversible

@@

header-includes:
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage[table]{xcolor}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{setspace}\singlespacing 
  - \newenvironment{zeroindent}
     {\par\setlength{\parindent}{0pt}}
     {\par}  
  - \setlength{\parskip}{0.25cm plus4mm minus3mm}

---
```{r setup, include=FALSE, eval=TRUE,echo=FALSE}
library("dplyr")
library("afex")
library("papaja")
library("magrittr")
library("kableExtra")

#As of 2018/12/07 apa_print() requires development version of Papaja
#devtools::install_github("crsh/papaja")
library(papaja)
```

Selective attention is commonly investigated using interference paradigms such as the Stroop [-@stroop_studies_1935] and flanker [@eriksen_effects_1974] tasks, where participants identify a target while ignoring a response-congruent or -incongruent distractor. Performance is typically better on congruent versus incongruent trials and the difference-the congruency effect-taken as an index of attentional priorities. Large congruency effects are thought to reflect ineffective filtering of the distracting stimuli whereas small congruency effects are thought to reflect effective filtering. By probing factors that systematically alter congruency effects, we can then make inferences about processes that control attentional filtering. For example, manipulating the frequency of conflict via the proportion of congruent versus incongruent trials has shown to influence the size of the congruency effect. Typically, a high proportion congruent experiment produces large congruency effects, whereas a low proportion congruent experiment produces small congruency effects [@logan_when_1979; @lowe_selective_1982; @west_effects_1998]. This result is usually  explained as strategic control, where participants increase attentional control under high-conflict demands and relax attentional control under low-conflict demands [@logan_attention_1980; @logan_strategies_1984; @logan_when_1979; @lowe_selective_1982]. Recent work however, has demonstrated that attentional control is not only adjusted by top-down regulation, but can also be triggered automatically by environmental cues [@brosowsky_memory-guided_2018; @bugg_support_2012; @egner_creatures_2014; @fischer_predicting_2015; @king_priming_2012; @mayr_outsourcing_2007].

For example, Crump, Gong, and Milliken [@crump_context-specific_2006; see also, @corballis_independent_2003] presented Stroop stimuli in one of two randomly chosen locations and manipulated the frequency of conflict associated with each location. One location was associated with a high frequency of conflict (25% congruent trials) and the other with a low frequency of conflict (75% congruent trials). Overall, the proportion of congruent trials was 50% and randomized such that the upcoming location could not be predicted. Even so, congruency effects were shown to be smaller for trials where the stimulus appeared in the high conflict location as compared to the low conflict location. This effect, known now as the context-specific proportion congruent effect (CSPC), has now been replicated in a number of different selective attention paradigms [e.g., @alards-tomalin_auditory_2017; @blais_trial-by-trial_2015; @bugg_conflict-triggered_2014; @crump_context-dependent_2018; @crump_learning_2016; @fischer_context-sensitive_2014; @hubner_location-specific_2016].

Critical evidence however, that CSPC effects reflect context-specific control rather than other non-control learning processes [e.g., @schmidt_stroop_2008], comes from work showing that CSPC effects can transfer to frequency unbiased items [@brosowsky_context-specific_2016; @crump_flexibility_2009; @weidler_attentional_2018; @weidler_transfer_2016; though, see @hutcheon_limits_2017]. Crump and Milliken (2009), for example, divided Stroop items into two mutually exclusive sets (e.g., RED/GREEN and BLUE/YELLOW). One set was defined as the frequency biased set, and presented with 75% congruency in one location, and 25% congruency in the other. The second set however, was presented with 50% congruency in both locations. Nevertheless, they found smaller congruency effects for unbiased items presented in the high conflict location as compared to the low conflict location.

One explanation for CSPC effects is that the repeated application of attentional priorities in a particular context creates an associative link in episodic memory between the attentional control procedures and the contextual information [@abrahamse_grounding_2016; @brosowsky_memory-guided_2018; @crump_learning_2016; @egner_creatures_2014]. Once the associative link is established, processing the context is assumed to trigger the retrieval of previous experiences, automatically reinstating the associated attentional priorities. Although this explanation seems plausible, it remains largely unclear how such associations are established and perhaps more importantly, it remains unclear what determines which learned associations will be used to guide attention at any given moment.

@@

Task-relevance might be one factor that is important for establishing and using context-attention associations. This proposal might seem counter-intuitive as it is typically assumed that the context, within selective attention tasks at least, is completely irrelevant to the task. In fact, referring to such stimulus dimensions as the "context" carries with it an implication that the feature is background information, irrelevant to the task-at-hand, and should be ignored by the participant. A closer look at CSPC designs however, suggests that the context cues may not be as irrelevant as implied. For example, the most commonly used context cue is location [@brosowsky_context-specific_2016; @corballis_independent_2003; @crump_context-specific_2006; @crump_context-specific_2008; @crump_learning_2016; @weidler_attentional_2018; @weidler_transfer_2016]. Although in some sense, the location is irrelevant for identifying the color of a word, it seems disingenuous to consider it completely task-irrelevant. In order to identify the target, the participant must first identify and orient to its location on the screen. In this is sense, it is relevant to their ability to complete the task and furthermore, location information may receive priority during encoding [@logan_what_1998; @mayr_spatial_1996]. Similar arguments could be made about other context features that are inherent to processing the target stimulus, like font-type [@bugg_multiple_2008] and color [@vietze_context_2009].

To date, we are aware of only one study that has looked at task-relevance within the CSPC literature. Crump, Vaquero, & Milliken [-@crump_context-specific_2008] used a Stroop prime-probe task where the distractor word is presented first followed by a color patch probe. Critically, they used the shape of the color patch probe, either a square or a circle, as the contextual dimension; one associated with a high frequency of conflict, the other with a low frequency. In one experiment participants were made aware of the shape-proportion contingencies but given no additional instructions. In the second, participants were given a secondary task to count the number of trials that contained a square. The first experiment failed to find evidence for context-specificity, whereas the second did. Although not testing task-relevance directly, @canadas_social_2013] found similar effects. In this study they used images of male and female faces as context cues in a flanker task. When given instructions to think of the faces as members of the gender categories they found CSPC effects, but when given instructions to think of the faces as individuals they did not. These results suggest that making a context dimension relevant to the on-going task is important for producing CSPC effects.

# The Current Study

Prior work suggests that the task-relevance of contextual information is important for establishing new associations between context cues and attentional priorities. Simply put, CSPC effects emerge when context cues are somehow made relevant to the task, but do not when made irrelevant [@canadas_social_2013; @crump_context-specific_2006; @crump_context-specific_2008]. However, it still remains unclear how task-relevance may or may not inform the use of previously learned associations. In the current study, we examined whether manipulating task-relevance would cause participants to ignore a previously learned association and adopt a new context-specific control strategy.

A flanker task was used to measure attentional control in an adapted CSPC design [@crump_flexibility_2009; @crump_reproducing_2017]. Similar to previous studies, flanker stimuli were presented in contexts associated with differing levels of conflict (0%, 50%, or 100% congruent). However, unlike previous CSPC designs where contexts were defined by a single discriminating feature (e.g., upper vs. lower screen locations), contexts were defined by two feature dimensions (see Figure 1A). In one condition, for example, contexts were defined by object identity (hat or chair) and color (blue or green). Only three out of the four possible feature combinations were presented, each associated with a different proportion of congruent trials (0%, 50%, or 100% congruent). Critically, the frequency unbiased context (50% congruent) always shared a feature with each of the frequency-biased contexts (0% and 100% congruent). The frequency biased contexts (0% and 100% congruent) however, did not share any features (see Figure 1C for an example).

A secondary counting task was used to manipulate the task-relevance of the context dimensions [@crump_flexibility_2009] which critically, switched halfway through the experiment. Participants were instructed to keep a running count of one of the overlapping features (e.g., "count whenever a hat is presented") which was associated with either a high or low frequency of conflict. Halfway through the experiment they received new instructions to count the other overlapping feature (e.g., "count whenever a green item is presented"). Critically, the set of context images remained the same throughout the experiment (see Figure 1C for an example) and the task-relevance of the context dimensions was the only aspect of the experiment that changed from the first to second phases.

The critical measure of interest is the congruency effect produced in the frequency unbiased context. If, on the one hand, task-relevance has no impact on the use of context-attention associations we would expect no differences between the congruency effects when the high-conflict context dimension is made task-relevant or the low-conflict context dimension is made task-relevant. This could occur because all three contexts are treated as individual contexts throughout the whole experiment [e.g., @canadas_social_2013]. Or it could occur because associations formed in the first block interfere with learning new associations in the second [e.g., @brosowsky_context-specific_2016]. On other hand, task-relevance might not only dictate how associations are formed, but also whether learned associations are used. In this case, we would expect smaller congruency effects when the high-conflict context dimension is made task-relevant as compared to when the low-conflict context dimension is made task-relevant. This would demonstrate that participants were able to ignore a previously learned association and adopt a new context-specific control strategy.

A secondary goal was to conceptually replicate previous findings and test the generalizability of any task-relevance effects. Therefore, we included three conditions that differed only in the kinds of stimuli used to create context cues. In one condition we used face images where context dimensions were defined by social categories, similar to @canadas_social_2013. However, they found that social categorization occurred spontaneously and it was unclear what impact that would have on the current results. Therefore, we included another condition that used images of simple objects where context dimensions were defined by object features. Finally, @canadas_social_2013 also found that context effects generalized to novel face images. In our third condition we used non-repeating face exemplars to examine whether any task-relevance effects would also generalize in our design or if learning would be image-specific.

# Methods

@@

## Participants

All participants were recruited from Amazon Mechanical Turk (AMT) and compensated $2.00 for participating. The amount compensated was calculated by estimating the maximum amount of time required to complete each experiment and multiplying by $6.00 per hour. For each experiment the number of HITs (Human intelligence tasks, an Amazon term for a work-unit) refers to the number of participants who initiated the study. Participants were included in the study if they completed all trials. We posted 150 HITs and 144 participants completed all trials.

## Apparatus & Stimuli

The experiments were programmed using JavaScript, CSS and HTML. The program allowed participants to complete task only if they were running Safari, Google Chrome, or Firefox web browsers. Flanker stimuli consisted of images of five arrows pointed left or right presented at 250 x 50 pixels (each arrow was 50 x 50 pixels). Context stimuli were constructed using images selected from @brady_visual_2013 color-rotated to blue and green, and face images from the Chicago Face Database [@ma_chicago_2015], supplemented with the NimStim Set of Facial Expressions [@tottenham_nimstim_2009]. The object images were displayed at 250 x 250 pixels, while the face images were displayed at 250 x 313 pixels. The experiment ran as a pop-up window that filled the entire screen. The background was white, and stimuli were presented in the center of the screen.

@@

## Procedure

@@

@@

# Results

Participants with mean error rates greater than 25% were excluded from the analyses, eliminating `r N_subjects_removed` participants. For all remaining participants, the correct RTs from frequency unbiased trials in each condition were submitted to an outlier removal procedure. The non-recursive Van Selst and Jolicoeur outlier removal procedure was applied after removing response times greater than 3000 ms [@van_selst_solution_1994]. This procedure removed `r round(percent_removed*100,digits = 2)`% of the total observations.

@@

The results of the RT analysis revealed the critical, two-way interaction between task-relevant context and unbiased-item congruency to be significant, `r RT_ANOVA$Task_Relevant_Context_Congruency`, demonstrating smaller congruency effects when the context dimension associated with high conflict was made task-relevant. Additionally, the three-way interaction between the task-relevant context, unbiased-item congruency, and context-type was non-significant, `r RT_ANOVA$Condition_Task_Relevant_Context_Congruency`, showing no evidence for differences between the CSPC effects across context-type manipulations.

Similarly, the corresponding error analysis also resulted in a significant two-way interaction between the task-relevant context and unbiased-item congruency, `r ACC_ANOVA$Task_Relevant_Context_Congruency`, and a non-significant three-way interaction between task-relevant context, unbiased-item congruency, and context-type, `r ACC_ANOVA$Condition_Task_Relevant_Context_Congruency`. Therefore, the congruency effects in error rates were also smaller when the context dimension associated with high conflict was made task-relevant, corroborating the results of the reaction time analysis.

\begingroup
\setlength{\parindent}{0in}
\vspace{2.5mm}
\textbf{Table 1. \\ \textit{Reaction times and error rates from Experiment 1.}}

![ ](images/table1.pdf) 
\endgroup

# General Discussion

The aim of the current study was to determine whether task-relevance plays a role in the contextual recruitment of selective attention. Specifically, we tested whether manipulating the relative task-relevance of context cues could cause participants to ignore a previously learned context-association and apply a new association. We created a frequency unbiased context cue that shared features with two frequency biased contexts and used a feature-counting task to manipulate the task-relevance of context dimensions across two blocks of trials. Critically, halfway through the experiment participants received new instructions changing the task-relevant feature from one frequency biased cue to the other.

The key finding was that the congruency effects for the frequency unbiased items were significantly larger when the low-conflict context was made task-relevant as compared to when the high-conflict context was made task-relevant. This result is consistent with prior CSPC effects and, like the previous work, suggests that the context cues triggered rapid adjustments to attentional control [@crump_flexibility_2009]. However, unlike prior studies, we were able to experimentally manipulate the CSPC effect across blocks of trials without changing any of the physical properties of the stimuli. This novel finding demonstrates that participants were able to learn and apply one context-attention association in the first phase, and subsequently ignore that association to learn a new association in the second phase.

This result implicates an important role for task-relevance in producing CSPC phenomena. @crump_context-specific_2008 used shapes as context cues in a prime-probe Stroop task and did not find CSPC effects until the context cues were made task-relevant. Similarly, @canadas_social_2013 eliminated the CSPC effect by making the contextual cue effectively unrelated to the task. These studies suggest that task-relevance plays an important role in establishing associations between contextual information and attentional priorities to produce CSPC effects. Our finding extends this work in two important ways. First, we show that changing the task-relevance of the presented cues corresponded with a change in attentional control in the predicted direction. This demonstrates that task-relevance is also a key determinant of whether a previously learned attention-context association will be used or ignored in favor of a new association. Second, we show that task-relevance allowed participants resolve competition between two competing contextual cues, responding on the basis of one at the expense of the other. To our knowledge, this is the first demonstration that CSPC effects can be produced when there are multiple, overlapping contextual cues available.

In light of prior work, we take this as evidence that the contextual recruitment of selective attention, although likely implicit, is not obligatory [e.g., @brosowsky_context-specific_2016], requiring that environmental information be incorporated into the task representation. Similarly, there appears to be flexibility in which environmental features are selected and used to guide attention, which can be rapidly updated depending on the task-relevance of those features. Such a result lends some insight into how context-specific control might operate in more complex, real-world environments, where there is an over-abundance of environmental features that afford many different learned associations. From a theoretical perspective, this result is consistent memory-based accounts of CSPC phenomena [@brosowsky_memory-guided_2018; @bugg_converging_2013; @crump_context-dependent_2018; @crump_flexibility_2009; @crump_reproducing_2017]. Under this view, a memory process encodes attentional priorities in the representation of individual experiences and, as a result, become associated with the environment where they were used. The subsequent reoccurrence of a prior context triggers the retrieval and reinstatement of those attentional priorities. Our results show however, that all the features of the environment may not be treated equally and that only task-relevant features are used to probe memory and guide attention.

@@

Traditional models of person perception posit that social categories are automatically activated in the presence of social stimuli  [e.g., @brewer_dual_1988; @devine_stereotypes_1989; @fiske_continuum_1990]. @canadas_social_2013 however, found that directing participants to think about faces in terms of individual features eliminated context-specific attention effect and suggested that momentary motivations may influence the automaticity of social categorization. Our results add to this literature by observing the influence of momentary motivations (i.e., task-relevance) when there is competition between two salient social categories. Specifically, we found that participants could categorize on the basis of one social cue at the expense of the other, and subsequently switch between them. These findings may speak to issues of automaticity in social categorization [@macrae_social_2001] as well as understanding how the situational context can prime one social identity over another [@crisp_multiple_2007]. Furthermore, we found no evidence for differences between the social and object conditions. This suggests that categorization within the CSPC task is quickly and easily learned when the task supports such learning and is non-unique to social stimuli.

>>>>

output:
#  papaja::apa6_word: default
#  papaja::apa6_pdf: default
   redoc::rdocx_reversible

@@

header-includes:
  - \usepackage{setspace}\singlespacing 
  - \AtBeginEnvironment{tabular}{\onehalfspacing}
  - \AtBeginEnvironment{lltable}{\onehalfspacing}
  - \captionsetup[table]{font={stretch=1.5}}
  - \captionsetup[figure]{font={stretch=1}}
  - \setlength{\skip\footins}{1cm}
  - \setlength{\footnotesep}{0.5cm}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage[table]{xcolor}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{setspace}\singlespacing 
  - \AtBeginEnvironment{tabular}{\onehalfspacing}
  - \AtBeginEnvironment{lltable}{\onehalfspacing}
  - \newenvironment{zeroindent}
     {\par\setlength{\parindent}{0pt}}
     {\par}

---

```{r setup, include=FALSE, eval=TRUE,echo=FALSE}
library("dplyr")
library("afex")
library("papaja")
library("magrittr")
library("kableExtra")
#As of 2018/12/07 apa_print() requires development version of Papaja
#devtools::install_github("crsh/papaja")
library(papaja)
```

Selective attention is commonly investigated using interference paradigms such
as the Stroop (1935) and flanker (Eriksen and Eriksen 1974) tasks, where
participants identify a target while ignoring a response-congruent or
-incongruent distractor. {++ Typically, p ++}{-- P --}erformance is
{-- typically --} better {++ (quicker reaction times and higher accuracy) ++} on
congruent versus incongruent trials and the difference-the congruency
effect-taken as an index of attentional priorities. Large congruency effects are
thought to reflect ineffective filtering of the distracting stimuli whereas
small congruency effects are thought to reflect effective filtering. By probing
factors that systematically alter congruency effects, we can then make
inferences about processes that control attentional filtering. For example,
manipulating the frequency of conflict via the proportion of congruent versus
incongruent trials has shown to influence the size of the congruency effect.
Typically, a high proportion congruent experiment produces large congruency
effects, whereas a low proportion congruent experiment produces small congruency
effects (Logan and Zbrodoff 1979; Lowe and Mitterer 1982; West and Baylis 1998).
This result is usually explained as strategic control, where participants
increase attentional control under high-conflict demands and relax attentional
control under low-conflict demands (Logan 1980; Logan, Zbrodoff, and Williamson
1984; Logan and Zbrodoff 1979; Lowe and Mitterer 1982). Recent work however, has
demonstrated that attentional control is not only adjusted by top-down
regulation, but can also be triggered automatically by environmental cues
(Brosowsky and Crump 2018; Bugg and Crump 2012; Egner 2014; Fischer and
Dreisbach 2015; King, Korb, and Egner 2012; Mayr and Bryck 2007).

For example, Crump, Gong, and Milliken (Crump, Gong, and Milliken 2006; see
also, Corballis and Gratton 2003) presented Stroop stimuli in one of two
randomly chosen locations and manipulated the frequency of conflict associated
with each location. One location was associated with a high frequency of
conflict (25% congruent trials) and the other with a low frequency of conflict
(75% congruent trials). Overall, the proportion of congruent trials was 50% and
randomized such that the upcoming location could not be predicted. Even so,
congruency effects were shown to be smaller for trials where the stimulus
appeared in the high conflict location as compared to the low conflict location.
This effect, known now as the context-specific proportion congruent effect
(CSPC), has now been replicated in a number of different selective attention
paradigms (e.g., Alards-Tomalin, Brosowsky, and Mondor 2017; Blais et al. 2015;
Bugg 2014; Crump et al. 2018; Crump 2016; Fischer, Gottschalk, and Dreisbach
2014; Hübner and Mishra 2016).

Critical evidence however, that CSPC effects reflect context-specific control
rather than other non-control learning processes (e.g., Schmidt and Besner
2008), comes from work showing that CSPC effects can transfer to frequency
unbiased items (Brosowsky and Crump 2016; Crump and Milliken 2009; Weidler, Dey,
and Bugg 2018; Weidler and Bugg 2016; though, see Hutcheon and Spieler 2017).
Crump and Milliken (2009), for example, divided Stroop items into two mutually
exclusive sets (e.g., RED/GREEN and BLUE/YELLOW). One set was defined as the
frequency biased set, and presented with 75% congruency in one location, and 25%
congruency in the other. The second set however, was presented with 50%
congruency in both locations. Nevertheless, they found smaller congruency
effects for unbiased items presented in the high conflict location as compared
to the low conflict location.

One explanation for CSPC effects is that the repeated application of attentional
priorities in a particular context creates an associative link in episodic
memory between the attentional control procedures and the contextual information
(Abrahamse et al. 2016; Brosowsky and Crump 2018; Crump 2016; Egner 2014). Once
the associative link is established, processing the context is assumed to
trigger the retrieval of previous experiences, automatically reinstating the
associated attentional priorities. Although this explanation seems plausible, it
remains largely unclear how such associations are established and perhaps more
importantly, it remains unclear what determines which learned associations will
be used to guide attention at any given moment.

@@

Task-relevance might be one factor that is important for establishing and using
context-attention associations. This proposal might seem counter-intuitive as it
is typically assumed that the context, within selective attention tasks at
least, is completely irrelevant to the task. In fact, referring to such stimulus
dimensions as the "context" carries with it an implication that the feature is
background information, irrelevant to the task-at-hand, and should be ignored by
the participant. A closer look at CSPC designs however, suggests that the
context cues may not be as irrelevant as implied. For example, the most commonly
used context cue is location (Brosowsky and Crump 2016; Corballis and Gratton
2003; Crump, Gong, and Milliken 2006; Crump, Vaquero, and Milliken 2008; Crump
2016; Weidler, Dey, and Bugg 2018; Weidler and Bugg 2016). Although in some
sense, the location is irrelevant for identifying the color of a word, it seems
disingenuous to consider it completely task-irrelevant. In order to identify the
target, the participant must first identify and orient to its location on the
screen. In this is sense, it is relevant to their ability to complete the task
and furthermore, location information may receive priority during encoding
(Logan 1998; Mayr 1996). Similar arguments could be made about other context
features that are inherent to processing the target stimulus, like font-type
(Bugg, Jacoby, and Toth 2008) and color (Vietze and Wendt 2009).

To date, we are aware of only one study that has looked at task-relevance within
the CSPC literature. Crump, Vaquero, & Milliken (2008) used a Stroop prime-probe
task where the distractor word is presented first followed by a color patch
probe. Critically, they used the shape of the color patch probe, either a square
or a circle, as the contextual dimension; one associated with a high frequency
of conflict, the other with a low frequency. In one experiment participants were
made aware of the shape-proportion contingencies but given no additional
instructions. In the second, participants were given a secondary task to count
the number of trials that contained a square. The first experiment failed to
find evidence for context-specificity, whereas the second did. Although not
testing task-relevance directly, Cañadas et al. (2013)\] found similar effects.
In this study they used images of male and female faces as context cues in a
flanker task. When given instructions to think of the faces as members of the
gender categories they found CSPC effects, but when given instructions to think
of the faces as individuals they did not. These results suggest that making a
context dimension relevant to the on-going task is important for producing CSPC
effects.

The Current Study
=================

Prior work suggests that the task-relevance of contextual information is
important for establishing new associations between context cues and attentional
priorities. Simply put, CSPC effects emerge when context cues are somehow made
relevant to the task, but do not when made irrelevant (Cañadas et al. 2013;
Crump, Gong, and Milliken 2006; Crump, Vaquero, and Milliken 2008). However, it
still remains unclear how task-relevance may or may not inform the use of
previously learned associations. In the current study, we examined whether
manipulating task-relevance would cause participants to ignore a previously
learned association and adopt a new context-specific control strategy.

A flanker task was used to measure attentional control in an adapted CSPC design
(Crump and Milliken 2009; Crump, Brosowsky, and Milliken 2017). Similar to
previous studies, flanker stimuli were presented in contexts associated with
differing levels of conflict (0%, 50%, or 100% congruent). However, unlike
previous CSPC designs where contexts were defined by a single discriminating
feature (e.g., upper vs. lower screen locations), contexts were defined by two
feature dimensions (see Figure 1A). In one condition, for example, contexts were
defined by object identity (hat or chair) and color (blue or green). Only three
out of the four possible feature combinations were presented, each associated
with a different proportion of congruent trials (0%, 50%, or 100% congruent).
Critically, the frequency unbiased context (50% congruent) always shared a
feature with each of the frequency-biased contexts (0% and 100% congruent). The
frequency biased contexts (0% and 100% congruent) however, did not share any
features (see Figure 1C for an example).

A secondary counting task was used to manipulate the task-relevance of the
context dimensions (Crump and Milliken 2009) which critically, switched halfway
through the experiment. Participants were instructed to keep a running count of
one of the overlapping features (e.g., "count whenever a hat is presented")
which was associated with either a high or low frequency of conflict. Halfway
through the experiment they received new instructions to count the other
overlapping feature (e.g., "count whenever a green item is presented").
Critically, the set of context images remained the same throughout the
experiment (see Figure 1C for an example) and the task-relevance of the context
dimensions was the only aspect of the experiment that changed from the first to
second phases.

The critical measure of interest is the congruency effect produced in the
frequency unbiased context. If, on the one hand, task-relevance has no impact on
the use of context-attention associations we would expect no differences between
the congruency effects when the high-conflict context dimension is made
task-relevant or the low-conflict context dimension is made task-relevant. This
could occur because all three contexts are treated as individual contexts
throughout the whole experiment (e.g., Cañadas et al. 2013). Or it could occur
because associations formed in the first block interfere with learning new
associations in the second (e.g., Brosowsky and Crump 2016). On other hand,
task-relevance might not only dictate how associations are formed, but also
whether learned associations are used. In this case, we would expect smaller
congruency effects when the high-conflict context dimension is made
task-relevant as compared to when the low-conflict context dimension is made
task-relevant. This would demonstrate that participants were able to ignore a
previously learned association and adopt a new context-specific control
strategy.

A secondary goal was to conceptually replicate previous findings and test the
generalizability of any task-relevance effects. Therefore, we included three
conditions that differed only in the kinds of stimuli used to create context
cues. In one condition we used face images where context dimensions were defined
by social categories, similar to Cañadas et al. (2013). However, they found that
social categorization occurred spontaneously and it was unclear what impact that
would have on the current results. Therefore, we included another condition that
used images of simple objects where context dimensions were defined by object
features. Finally, Cañadas et al. (2013) also found that context effects
generalized to novel face images. In our third condition we used non-repeating
face exemplars to examine whether any task-relevance effects would also
generalize in our design or if learning would be image-specific.

Methods
=======

@@

Participants
------------

All participants were recruited from Amazon Mechanical Turk (AMT) and
compensated \$2.00 for participating. The amount compensated was calculated by
estimating the maximum amount of time required to complete each experiment and
multiplying by \$6.00 per hour. For each experiment the number of HITs (Human
intelligence tasks, an Amazon term for a work-unit) refers to the number of
participants who initiated the study. Participants were included in the study if
they completed all trials. We posted 150 HITs and 144 participants completed all
trials.

Apparatus & Stimuli
-------------------

The experiments were programmed using JavaScript, CSS and HTML. The program
allowed participants to complete task only if they were running Safari, Google
Chrome, or Firefox web browsers. Flanker stimuli consisted of images of five
arrows pointed left or right presented at 250 x 50 pixels (each arrow was 50 x
50 pixels). Context stimuli were constructed using images selected from Brady et
al. (2013) color-rotated to blue and green, and face images from the Chicago
Face Database (Ma, Correll, and Wittenbrink 2015), supplemented with the NimStim
Set of Facial Expressions (Tottenham et al. 2009). The object images were
displayed at 250 x 250 pixels, while the face images were displayed at 250 x 313
pixels. The experiment ran as a pop-up window that filled the entire screen. The
background was white, and stimuli were presented in the center of the screen.

@@

Procedure
---------

@@

Results
=======

@@

Participants with mean error rates greater than 25% were excluded from the
analyses, eliminating `r N_subjects_removed` participants. For all remaining participants,
the correct RTs from frequency unbiased trials in each condition were submitted
to an outlier removal procedure. The non-recursive Van Selst and Jolicoeur
outlier removal procedure was applied after removing response times greater than
3000 ms (Van Selst and Jolicoeur 1994). This procedure removed `r round(percent_removed*100,digits = 2)`% of
the total observations.

@@

The results of the RT analysis revealed the critical, two-way interaction
between task-relevant context and unbiased-item congruency to be significant,
$F(1,128) = 8.30$`r RT_ANOVA$Task_Relevant_Context_Congruency` $\text{MSE} = 1,608.54$`r RT_ANOVA$Task_Relevant_Context_Congruency`
$p = .005$`r RT_ANOVA$Task_Relevant_Context_Congruency` ${\widehat{\eta}}_{p}^{2} = .061$, demonstrating smaller
congruency effects when the context dimension associated with high conflict was
made task-relevant. Additionally, the three-way interaction between the
task-relevant context, unbiased-item congruency, and context-type was
non-significant, $F(2,128) = 0.00$`r RT_ANOVA$Condition_Task_Relevant_Context_Congruency`
$\text{MSE} = 1,608.54$`r RT_ANOVA$Condition_Task_Relevant_Context_Congruency` $p = .996$`r RT_ANOVA$Condition_Task_Relevant_Context_Congruency`
${\widehat{\eta}}_{p}^{2} = .000$, showing no evidence for differences between
the CSPC effects across context-type manipulations.

Similarly, the corresponding error analysis also resulted in a significant
two-way interaction between the task-relevant context and unbiased-item
congruency, $F(1,128) = 6.72$`r ACC_ANOVA$Task_Relevant_Context_Congruency` $\text{MSE} = 13.93$`r ACC_ANOVA$Task_Relevant_Context_Congruency`
$p = .011$`r ACC_ANOVA$Task_Relevant_Context_Congruency` ${\widehat{\eta}}_{p}^{2} = .050$, and a non-significant
three-way interaction between task-relevant context, unbiased-item congruency,
and context-type, $F(2,128) = 0.92$`r ACC_ANOVA$Condition_Task_Relevant_Context_Congruency` $\text{MSE} = 13.93$`r ACC_ANOVA$Condition_Task_Relevant_Context_Congruency`
$p = .400$`r ACC_ANOVA$Condition_Task_Relevant_Context_Congruency` ${\widehat{\eta}}_{p}^{2} = .014$. Therefore, the
congruency effects in error rates were also smaller when the context dimension
associated with high conflict was made task-relevant, corroborating the results
of the reaction time analysis.

General Discussion
==================

The aim of the current study was to investigate whether task-relevance plays a
role in the contextual recruitment of selective attention. Specifically, we
tested whether manipulating the relative task-relevance of context cues could
cause participants to ignore a previously learned context-association and apply
a new association. We created a frequency unbiased context cue that shared
features with two frequency biased contexts and included a feature-counting task
to manipulate the task-relevance of context dimensions across two blocks of
trials. Critically, halfway through the experiment participants received new
instructions changing the task-relevant feature from one frequency biased cue to
the other.

The key finding was that the congruency effects for the frequency unbiased items
were significantly larger when the low-conflict context was made task-relevant
as compared to when the high-conflict context was made task-relevant. This
result is consistent with prior CSPC effects and, like the previous work,
suggests that the context cues triggered rapid adjustments to attentional
control (Crump and Milliken 2009). However, unlike prior studies, we were able
to experimentally manipulate the CSPC effect across blocks of trials without
changing any of the physical properties of the stimuli. This novel finding
demonstrates that participants were able to learn and apply one
context-attention association in the first phase, and subsequently ignore that
association to learn a new association in the second phase.

This result implicates an important role for task-relevance in producing CSPC
phenomena. Crump, Vaquero, and Milliken (2008) used shapes as context cues in a
prime-probe Stroop task and did not find CSPC effects until the context cues
were made task-relevant. Similarly, Cañadas et al. (2013) eliminated the CSPC
effect by making the contextual cue effectively unrelated to the task. These
studies suggest that task-relevance plays an important role in establishing
associations between contextual information and attentional priorities to
produce CSPC effects. Our finding extends this work in two important ways.
First, we show that changing the task-relevance of the presented cues
corresponded with a change in attentional control in the predicted direction.
This demonstrates that task-relevance is also a key determinant of whether a
previously learned attention-context association will be used or ignored in
favor of a new association. Second, we show that task-relevance allowed
participants resolve competition between two competing contextual cues,
responding on the basis of one at the expense of the other. To our knowledge,
this is the first demonstration that CSPC effects can be produced when there are
multiple, overlapping contextual cues available.

In light of prior work, we take this as evidence that the contextual recruitment
of selective attention, although likely implicit, is not obligatory (e.g.,
Brosowsky and Crump 2016), requiring that environmental information be
incorporated into the task representation. Similarly, there appears to be
flexibility in which environmental features are selected and used to guide
attention, which can be rapidly updated depending on the task-relevance of those
features. Such a result lends some insight into how context-specific control
might operate in more complex, real-world environments, where there is an
over-abundance of environmental features that afford many different learned
associations. From a theoretical perspective, this result is consistent
memory-based accounts of CSPC phenomena (Brosowsky and Crump 2018; Bugg and
Hutchison 2013; Crump et al. 2018; Crump and Milliken 2009; Crump, Brosowsky,
and Milliken 2017). Under this view, a memory process encodes attentional
priorities in the representation of individual experiences and, as a result,
become associated with the environment where they were used. The subsequent
reoccurrence of a prior context triggers the retrieval and reinstatement of
those attentional priorities. Our results show however, that all the features of
the environment may not be treated equally and that only task-relevant features
are used to probe memory and guide attention.

@@

Traditional models of person perception posit that social categories are
automatically activated in the presence of social stimuli (e.g., Brewer 1988;
Devine 1989; Fiske and Neuberg 1990). Cañadas et al. (2013) however, found that
directing participants to think about faces in terms of individual features
eliminated context-specific attention effect and suggested that momentary
motivations may influence the automaticity of social categorization. Our results
add to this literature by observing the influence of momentary motivations
(i.e., task-relevance) when there is competition between two salient social
categories. Specifically, we found that participants could categorize on the
basis of one social cue at the expense of the other, and subsequently switch
between them. These findings may speak to issues of automaticity in social
categorization (Macrae and Bodenhausen 2001) as well as understanding how the
situational context can prime one social identity over another (Crisp and
Hewstone 2007). Furthermore, we found no evidence for differences between the
social and object conditions. This suggests that categorization within the CSPC
task is quickly and easily learned when the task supports such learning and is
non-unique to social stimuli.

In sum, our results provide new evidence that changes in task-relevance can
update the contextual recruitment of selective attention in a CSPC flanker task.
Importantly, we show that making one feature task-relevant produced a CSPC
effect, even in the presence of a competing contextual cue. Moreover, changing
the task-relevant context-cue midway through the experiment was accompanied by
predictable changes in the congruency effects. This effect was found to be
generalizable across two different kinds of stimuli and occurred even when using
non-repeating images. These results implicate an important role for
task-relevance in producing context-dependency and are consistent with
memory-based accounts of CSPC phenomena.

References
==========

Abrahamse, Elger, Senne Braem, Wim Notebaert, and Tom Verguts. 2016. "Grounding
Cognitive Control in Associative Learning." *Psychological Bulletin* 142:
693--728. <https://doi.org/10.1037/bul0000047>.
